\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{occi}
\setlength{\headheight}{13pt}
\pagestyle{fancy}

% default sans-serif
\renewcommand{\familydefault}{\sfdefault}

% no lines for headers and footers
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% header
\fancyhf{}
\lhead{GFD-P-R.185}
\rhead{\today}

% footer
\lfoot{occi-wg@ogf.org}
\rfoot{\thepage}

% paragraphs need some space...
\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

% some space between header and text...
\headsep 13pt

\setcounter{secnumdepth}{4}

\begin{document}

% header on first page is different
\thispagestyle{empty}

GFD-P-R.185 \hfill Thijs Metsch, Platform Computing\\
OCCI-WG \hfill Andy Edmonds, Intel\\
\rightline {October 7, 2010}\\
\rightline {Updated: \today}

\vspace*{0.5in}

\begin{Large}
\textbf{Open Cloud Computing Interface - RESTful HTTP Rendering}
\end{Large}

\vspace*{0.5in}

\underline{Status of this Document}

This document provides information to the community regarding the
specification of the Open Cloud Computing Interface. Distribution is
unlimited.

\underline{Copyright Notice}

Copyright \copyright ~Open Grid Forum (2009-2011). All Rights Reserved.

\underline{Trademarks}

OCCI is a trademark of the Open Grid Forum.

\underline{Abstract}

This document, part of a document series, produced by the OCCI working
group within the Open Grid Forum (OGF), provides a high-level
definition of a Protocol and API. The document is based upon
previously gathered requirements and focuses on the scope of important
capabilities required to support modern service offerings.

\newpage
\tableofcontents
\newpage

\section{Introduction}
\input{include/introduction}

\section{Notational Conventions}
\input{include/notational}

\todo{update rfc}

This document uses the Augmented Backus-Naur Form (ABNF) notation of
RFC 2616 \cite{rfc2616}, and explicitly includes the following rules
from it: quoted-string, token, SP (space), LOALPHA, DIGIT.

The following terms \cite{rfc3986} are used when referring to URI
components:

\begin{verbatim}
 http://example.com:8080/over/there?action=stop#xyz
 \__/   \______________/\_________/ \_________/ \_/
  |           |            |            |        |
scheme     authority       path        query   fragment
\end{verbatim}

\section{OCCI HTTP Rendering}
The OCCI HTTP Rendering document specifies how the OCCI Core Model
\cite{occi:core}, including extensions thereof, is rendered over the
HTTP protocol \cite{rfc2616}. 

\subsection{Introduction}
The OCCI HTTP Rendering uses many of the features the HTTP and
underlying protocols offer and builds upon the Resource Oriented
Architecture (ROA). ROA's use Representation State Transfer (REST)
\cite{REST_Fielding} to cater for client and service
interactions. Interaction with the system is by inspection and
modification of a set of related resources and their states, be it on
the complete state or a sub-set. Resources MUST be uniquely
identified. HTTP is an ideal protocol to use in ROA systems as it
provides the means to uniquely identify individual resources through
URLs as well as operating upon them with a set of general-purpose
methods known as HTTP verbs. These HTTP verbs map loosely to the
resource related operations of Create (POST), Retrieve (GET), Update
(POST/PUT) and Delete (DELETE).

Each resource instance within an OCCI system MUST have a unique
identifier stored in the \emph{occi.core.id} attribute of the
\hl{Entity} type \cite{occi:core}.
It is RECOMMENDED to use a Uniform Resource Name (URN) as the
identifier stored in \emph{occi.core.id}.

The structure of these identifiers is opaque and the system should not
assume a static, pre-determined scheme for their structure. For
example \emph{occi.core.id} could be
\emph{urn:uuid:de7335a7-07e0-4487-9cbd-ed51be7f2ce4}.

The following sections and paragraphs describe how the OCCI model MUST
be implemented by OCCI implementations. Operations which are not
defined are out of scope for this specification and MAY be implemented. 
This is the minimal set to ensure interoperability.

\subsection{Namespace}

The OCCI HTTP Protocol maps the OCCI Core model into the URL hierarchy by binding
\hl{Kind} and \hl{Mixin} instances to unique URL paths. Such a URL path is called
the {\em location} of the \hl{Kind} or \hl{Mixin}.
A provider is free to choose the {\em location} as long as it is unique
within the service provider's URL namespace. 

Location paths are defined through specific renderings and MUST be present for all \hl{Kind} or \hl{Mixin} instances. The location
paths MUST end with a \emph{'/'}. The {\em location} MAY be realized by applying URI templates defined by \cite{rfc6570}.

A \hl{Kind} instance whose associated type cannot be instantiated MUST NOT be
bound to an URL path. This applies to the \hl{Kind} instance for OCCI Entity
which, according to OCCI Core, cannot be instantiated \cite{occi:core}.

\subsubsection{unbound/bound}

Since a limited set of URL paths are bound to \hl{Kind} and \hl{Mixin}
instances the URL hierarchy consists of both {\em bound} and {\em unbound}
paths. A bound URL path is the {\em location} of a \hl{Kind} or \hl{Mixin} collection.

An unbound URL path MAY represent the union of all \hl{Kind} and \hl{Mixin}
collection ``below'' the unbound path.

\subsection{HTTP}

\subsubsection{General Status codes}

\subsubsection{Versioning}

\subsubsection{Media types}

\paragraph{text uri-list (rendering)}

\subsubsection{Well known QI}

\subsubsection{Async op}

\subsubsection{Import Export}

\begin{itemize}
\item realize as Category
\item HTTP POST user defined mixin
\item HTTP POST against that collection with Content-Type…
\item HTTP GET against that collection with Accept...
\end{itemize}

\subsection{Operations}

\subsubsection{Entity instances}
POST
partial-update
action
PUT
create
full-update
GET
retrieval
DELETE
delete
Link instances
POST
inline creation 
\subsubsection{Collections}
POST
create entity instance
add entity to mixin
action
PUT
full-update on mixins...not kinds.
full-update on kind collection -> not defined
GET + filter/query?
retrieval
DELETE
delete all or single (!) entity
remove entity of mixins 
\subsubsection{QI}
POST
user defined mixin
PUT
N/A
GET + filter/query!
retrieve capabilities
DELETE
Removal user defined mixin

\section{Security Considerations}
\label{sec:sec_consid}
The OCCI HTTP rendering assumes HTTP or HTTP-related mechanisms for
security. As such, implementations SHOULD support
TLS \footnote{http://datatracker.ietf.org/wg/tls/} for transport layer
security.

Authentication SHOULD be realized by HTTP authentication mechanisms,
namely HTTP Basic or Digest Auth \cite{rfc2617}, with the former as
default. Additional profiles MAY specify other methods and should
ensure that the selected authentication scheme can be renderable over
the HTTP or HTTP-related protocols.

Authorization is not enforced on the protocol level, but SHOULD be
performed by the implementation. For the authorization decision, the
authentication information as provided by the mechanisms described
above MUST be used.

Protection against potential Denial-of-Service scenarios are out of
scope of this document; the OCCI HTTP Rendering specifications assumes
cooperative clients that SHOULD use selection and filtering as
provided by the Category mechanism wherever possible. Additional
profiles to this document, however, MAY specifically address such
scenarios; in that case, best practices from the HTTP ecosystem and
appropriate mechanisms as part of the HTTP protocol specification
SHOULD be preferred.

As long as specific extensions of the OCCI Core and Model
specification do not impose additional security requirements than the
OCCI Core and Model specification itself, the security considerations
documented above apply to all (existing and future)
extensions. Otherwise, an additional profile to this specification
MUST be provided; this profile MUST express all additional security
considerations using HTTP mechanisms.

\section{Glossary}
\label{sec:glossary}
\input{include/glossary}
 
\section{Contributors}
\input{include/contributors}

\section{Intellectual Property Statement}
\input{include/ip}

\section{Disclaimer}
\input{include/disclaimer}

\section{Full Copyright Notice}
\input{include/copyright}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}