\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[activate={true,nocompatibility},final,tracking=true,kerning=true,spacing=true]{microtype}
\usepackage[plainpages=false,pdfpagelabels,unicode]{hyperref}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{comment}
\usepackage{occi}
\usepackage{lineno}   % adds line numbers, may be removed for non draft versions
\linenumbers          % adds line numbers, may be removed for non draft versions
\usepackage{verbatim} % adds verbatim options
\usepackage{tabularx} % adds extended tabular formatting options
\usepackage{listings}
\usepackage{color}
\usepackage{appendix}
\definecolor{lightgray}{rgb}{.9,.9,.9}
\definecolor{darkgray}{rgb}{.4,.4,.4}
\definecolor{purple}{rgb}{0.65, 0.12, 0.82}

\lstdefinelanguage{json}{
  ndkeywords={String, Number, Boolean, Null, Object, Array},
  ndkeywordstyle=\itshape
}
\lstset{
   language=json,
   basicstyle=\footnotesize,
   xleftmargin=1.5cm
}

\setlength{\headheight}{13pt}
\pagestyle{fancy}

% default sans-serif
\renewcommand{\familydefault}{\sfdefault}

% no lines for headers and footers
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% header
\fancyhf{}
\lhead{GWD-R}
\rhead{\today}

% footer
\lfoot{occi-wg@ogf.org}
\rfoot{\thepage}

% paragraphs need some space...
\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

% some space between header and text...
\headsep 13pt

\setcounter{secnumdepth}{4}

\begin{document}

% header on first page is different
\thispagestyle{empty}

Draft \hfill Ralf Nyrén, Independent \\
OCCI-WG \hfill Florian Feldhaus, Independent\\
\rightline {Boris Parák, CESNET}\\
\rightline {Zden\v{e}k \v{S}ustr, CESNET}\\
\rightline {February 25, 2011}\\
\rightline {Updated: \today}

\vspace*{0.5in}

\begin{Large}
\textbf{Open Cloud Computing Interface -- JSON Rendering}
\end{Large}

\vspace*{0.5in}

\underline{Status of this Document}

\input{include/status}

\underline{Copyright Notice}

Copyright \copyright Open Grid Forum (2012-2016). All Rights Reserved.

\underline{Trademarks}

OCCI is a trademark of the Open Grid Forum.

\underline{Abstract}

\input{include/abstract}

\newpage
\tableofcontents
\newpage

\section{Introduction}
\input{include/introduction}

\section{Notational Conventions}
\input{include/notational}

\section{OCCI JSON Rendering}
\label{sec:json_format}
The OCCI JSON Rendering specifies a rendering of OCCI instance types in the JSON
data interchange format as defined in \cite{rfc4627}.

The rendering can be used to render OCCI instances independently of the
protocol being used. Thus messages can be delivered by, e.g., the HTTP
protocol as specified in \cite{occi:http_protocol}.

The following media-type MUST be used for the OCCI JSON Rendering:

{\tt application/occi+json}

The OCCI JSON Rendering consists of a JSON object containing information on the
OCCI Core instances OCCI Kind, OCCI Mixin, OCCI Action,
OCCI Link and OCCI Resource. The rendering also include a JSON object to invoke
the operation identified by OCCI Actions.
The rendering of each OCCI Core instance will be
described in the following sections.

\subsection{Entity Instance Rendering}
\label{sec:format_entity_instance_rendering}

Entity instances MUST be rendered as JSON hashmaps.

\subsubsection{Resource Instance Rendering}
\label{sec:format_resource}

The OCCI Resource Instance Rendering consists of a JSON object as shown in the
following declaration. Appendix~\ref{resouce_instance_rendering_example} contains a detailed
example.
Table~\ref{tbl:format_resource} defines the object members.
\begin{lstlisting}
{
    "kind": String,
    "mixins": Array,
    "attributes": Object,
    "actions": Array,
    "id": String,
    "links": Array,
    "summary": String,
    "title": String,
}
\end{lstlisting}


\mytablefloat{
    \label{tbl:format_resource}
    OCCI Resource instance rendered with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llllX}
    \toprule
    Object member & JSON type         & Mutability  & Multiplicity  & Description  \\
    \colrule
    kind          & String            & immutable   & 1             & Type identifier. \\

    mixins        & Array of Strings  & mutable     & 0..1          & List of type identifiers of associated OCCI Mixins.  \\

    attributes    & Object            & mutable     & 0..1          & Instance Attributes (see \ref{sec:format_attribute_description}). \\

    actions       & Array of Strings  & mutable     & 0..1          & List of type identifiers of OCCI Actions applicable to the OCCI Resource instance. \\

    id            & String            & immutable   & 1             & ID of the OCCI Resource. \\

    links         & Array of Objects  & mutable     & 0..1          & List of OCCI Links (fully rendered instances, see \ref{sec:format_link}). \\

    summary       & String            & mutable     & 0..1          & Summary text of resource. \\

    title         & String            & mutable     & 0..1          & Title of resource. \\
    \botrule
    \end{tabularx}
}

\paragraph{Action Invocation Rendering}
\label{sec:format_action_invocation}

The OCCI Action Invocation Rendering identifies an invocable operation on a OCCI Resource or
OCCI Link instance. To trigger such an operation the OCCI Action Invocation
Rendering is required.

The OCCI Action Invocation Rendering consists of a top-level JSON object as shown in the
following declaration. Appendix~\ref{action_invocation_rendering_example} contains a detailed example.
Table~\ref{tbl:format_action_invocation} defines the object members.

\begin{lstlisting}
{
    "action": String,
    "attributes": Object
}
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_action_invocation}
    An OCCI Action invocation is rendered with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llllX}
    \toprule
    Object member   & JSON type   & Mutability  & Multiplicity  & Description \\
    \colrule
    action          & String      & immutable   & 1             & Type identifier. \\

    attributes      & Object      & mutable     & 0..1          & Instance attributes (see \ref{sec:format_attribute_description}). \\
    \botrule
    \end{tabularx}
}


\subsubsection{Link Instance Rendering}
\label{sec:format_link}

The OCCI Link Instance Rendering consists of a JSON object as shown in the
following declaration. Appendix~\ref{link_instance_rendering_example} contains a detailed example.
Table~\ref{tbl:format_link} defines the object members.
\begin{lstlisting}
{
    "kind": String,
    "mixins": Array,
    "attributes": Object,
    "actions": Array,
    "id": String,
    "source": Object,
    "target": Object,
    "title": String
}
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_link}
    OCCI Link instances are rendered with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llllX}
    \toprule
    Object member & JSON type         & Mutability  & Multiplicity  & Description \\
    \colrule
    kind          & String            & immutable   & 1             & Type identifier. \\

    mixins        & Array of Strings  & mutable     & 0..1          & List of type identifiers of associated OCCI Mixins. \\

    attributes    & Object            & mutable     & 0..1          & Instance attributes (see \ref{sec:format_attribute_description}). \\

    actions       & Array of Strings  & mutable     & 0..1          & List of type identifiers of OCCI Action Categories applicable to the OCCI Link instance. \\

    id            & String            & immutable   & 1             & ID of the OCCI Link.\\

    source        & Object            & immutable   & 1             & Hashmap of the link source (see \ref{par:format_link_src_trgt}). \\

    target        & Object            & immutable   & 1             & Hashmap of the link target (see \ref{par:format_link_src_trgt}). \\

    title         & String            & mutable     & 0..1          & Title of the Link \\
    \botrule
    \end{tabularx}
}

\paragraph{Link Instance Source/Target Rendering}
\label{par:format_link_src_trgt}\hfill\\

The OCCI Link Instance Source/Target Rendering consists of a JSON object as shown in the
following declaration. Appendix~\ref{link_instance_rendering_example} contains a detailed example.
Table~\ref{tbl:format_link_src_trgt} defines the object members. {\tt location} maps to OCCI
Core's {\tt source} and {\tt target} model attributes and {\tt kind} maps to OCCI Core's {\tt target.kind}
model attribute. The value of {\tt kind} for {\tt source} is implied by OCCI Core's model attribute value
for {\tt source}.
\begin{lstlisting}
{
    "location": String,
    "kind": String
}
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_link_src_trgt}
    OCCI Link sources/targets are rendered with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llllX}
    \toprule
    Object member & JSON type         & Mutability  & Multiplicity  & Description \\
    \colrule
    location      & String            & immutable   & 1             & URI of the link target/source. \\

    kind          & String            & immutable   & 0..1          & Kind identifier, supplied if location points to an OCCI Resource. \\
    \botrule
    \end{tabularx}
}

\subsection{Category Instance Rendering}
\label{sec:format_category_instance_rendering}
Category instances MUST be rendered as JSON hashmaps.

\subsubsection{Kind Instance Rendering}
\label{sec:format_kind}

The OCCI Kind Instance Rendering consists of a JSON object as shown in the
following declaration. Appendix~\ref{kind_instance_rendering_example} contains a detailed example.
Table~\ref{tbl:format_kind} defines the top-level object members.

\mytablefloat{
    \label{tbl:format_kind}
    OCCI Kind instances are rendered with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llllX}
    \toprule
    Object member   & JSON type         & Mutability  & Multiplicity  & Description \\
    \colrule
    term            & String            & immutable   & 1             & Unique identifier within the categorization scheme. \\

    scheme          & String            & immutable   & 1             & Categorization scheme. \\

    title           & String            & immutable   & 0..1          & Title of the OCCI Kind. \\

    attributes      & Object            & immutable   & 0..1          & Attribute description, see~\ref{tbl:format_attribute_description}. \\

    parent          & String            & immutable   & 0..1          & OCCI Kind type identifier of the related ``parent'' \hl{Kind} instance. \\

    actions         & Array of Strings  & immutable   & 0..1          & List of OCCI Action type identifiers. \\

    location        & String            & immutable   & 0..1          & Transport protocol specific URI bound to the OCCI Kind instance. MUST be supplied for the OCCI Kinds of all OCCI Entities except OCCI Entity itself. \\
    \botrule
    \end{tabularx}
}

\begin{lstlisting}
{
    "term": String,
    "scheme": String,
    "title": String,
    "attributes": Object,
    "actions": Array,
    "parent": String,
    "location": String
}
\end{lstlisting}

\subsubsection{Mixin Instance Rendering}
\label{sec:format_mixin}

The OCCI Mixin Instance Rendering consists of a JSON object as shown in the following
declaration. Appendix~\ref{mixin_instance_rendering_example} contains a detailed example.
Table~\ref{tbl:format_mixin} defines the top-level object members.

\begin{lstlisting}
{
    "term": String,
    "scheme": String,
    "title": String,
    "attributes": Object,
    "actions": Array,
    "depends": Array,
    "applies": Array,
    "location": String
}
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_mixin}
    OCCI Mixin instances are rendered with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llllX}
    \toprule
    Object member   & JSON type         & Mutability  & Multiplicity  & Description \\
    \colrule
    term            & String            & immutable   & 1             & Unique identifier within the categorization scheme. \\

    scheme          & String            & immutable   & 1             & Categorization scheme. \\

    title           & String            & immutable   & 0..1          & Title of the OCCI Mixin. \\

    attributes      & Object            & immutable   & 0..1          & Attribute description, see~\ref{tbl:format_attribute_description}. \\

    depends         & Array of Strings  & immutable   & 0..1          & List of type identifiers of the dependent \hl{Mixin} instances. \\

    applies         & Array of Strings  & immutable   & 0..1          & List of OCCI Kind type identifiers this OCCI Mixin can be applied to. \\

    actions         & Array of Strings  & immutable   & 0..1          & List of OCCI Action type identifiers. \\

    location        & String            & immutable   & 1             & Transport protocol specific URI bound to the OCCI Mixin instance. \\
    \botrule
    \end{tabularx}
}

\subsubsection{Action Instance Rendering}
\label{sec:format_action}

The OCCI Action Instance Rendering consists of a JSON object as shown in the
following declaration. Appendix~\ref{action_instance_rendering_example} contains a detailed example.
Table~\ref{tbl:format_action} defines the top-level object members.

\mytablefloat{
    \label{tbl:format_action}
    OCCI Actions are rendered inside the top-level JSON object with name {\em actions} as an array of JSON Objects with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llllX}
    \toprule
    Object member   & JSON type   & Mutability  & Multiplicity  & Description \\
    \colrule
    term            & String      & immutable   & 1             & Unique type identifier within the categorization scheme. \\

    scheme          & String      & immutable   & 1             & Categorization scheme. \\

    title           & String      & immutable   & 0..1          & Title of the OCCI Action. \\

    attributes      & Object      & immutable   & 0..1          & Attribute description, see~\ref{tbl:format_attribute_description}. \\
    \botrule
    \end{tabularx}
}

\begin{lstlisting}
{
    "term": String,
    "scheme": String,
    "title": String,
    "attributes": Object,
}
\end{lstlisting}


\subsection{Entity Collection Rendering}
Collections of Entity instances MUST be rendered as JSON arrays. The content of that array is a set
of entity instance renderings.

That array MUST be a member of a JSON hashmap that is associated with the relevant key name specific
to the type of Entity collection being rendered.

\subsubsection{Resource Collection Rendering}

The JSON hashmap key-name associated with the array of resource instances MUST be \hl{resources}.

\begin{lstlisting}
{
    "resources": []
}
\end{lstlisting}

\subsubsection{Link Collection Rendering}

The JSON hashmap key-name associated with the array of link instances MUST be \hl{links}.

\begin{lstlisting}
{
    "links": []
}
\end{lstlisting}

\subsection{Category Collection Rendering}
Collections of Category instances MUST be rendered as JSON arrays. The content of that array
is a set of Category instance renderings.

That array MUST be a member of a JSON hashmap that is associated with the relevant key name
specific to the type of Category collection being rendered.


\subsubsection{Kind Collection Rendering}

The JSON hashmap key-name associated with the array of kind instances MUST be \hl{kinds}.

\begin{lstlisting}
{
    "kinds": []
}
\end{lstlisting}

\subsubsection{Mixin Collection Rendering}

The JSON hashmap key-name associated with the array of mixin instances MUST be \hl{mixins}.

\begin{lstlisting}
{
    "mixins": []
}
\end{lstlisting}

\subsubsection{Action Collection Rendering}

The JSON hashmap key-name associated with the array of action instances MUST be \hl{actions}.

\begin{lstlisting}
{
    "actions": []
}
\end{lstlisting}


Collections of Category instances are rendered as JSON arrays.

\subsection{Attributes Rendering}

Attribute names consist of alphanumeric characters separated by dots. The dots
define a logical namespace-like hierarchy. This hierarchy is NOT reflected in JSON
objects. As shown in the following declaration, the attribute name is an opaque
identifier rendered as hashmap \textit{key}. The hashmap \textit{value} contains either a
Number, a String, a Boolean, an Array or an Object (as an attribute value or an attribute
description, following the Attribute Description Rendering, see \ref{sec:format_attribute_description}).
\begin{lstlisting}
{
    "one.two.three": Number|String|Boolean|Array|Object,
    "one.two.four" : Number|String|Boolean|Array|Object
}
\end{lstlisting}

For examples of rendered Attributes please refer for instance to the Resource instance
rendering example in Appendix~\ref{resouce_instance_rendering_example}.


\subsubsection{Attribute Description Rendering}
\label{sec:format_attribute_description}

Attribute Descriptions are rendered as JSON objects as defined in table~\ref{tbl:format_attribute_description}

\mytablefloat{
    \label{tbl:format_attribute_description}
    All properties of the Attribute definition are optional, but may contain
defaults which MUST be used if the Attribute is not present in the instantiated
OCCI Entity.
    } {
    \begin{tabularx}{\textwidth}{lllX}
    \toprule
    Object member   & JSON type                               & Default   & Description \\
    \colrule
    mutable         & Boolean                                 & false     & Defines if the Attribute is mutable after initialization. \\

    required        & Boolean                                 & false     & Defines if the Attribute MUST be specified at instantiation of the OCCI Entity. \\

    type            & String                                  & string    & Type of the Attribute. MUST be either ``string'', ``number'', ``boolean'', ``array'' or ``object''. \\

    pattern         & Object                                  & {}        & JSON Schema \cite{json-schema} to validate the value of the attribute. It is recommended to specify the \texttt{\$schema} property for the schema used. \\

    default         & String, Number, Boolean, Array, Object  & {\tt (none)} & Attribute default. MUST be the same type as defined in the type property and MUST be used if the Attribute is not present in the instantiated OCCI Entity. \\

    description     & String                                  & {\tt (none)} & Description of the attribute. \\
    \botrule
    \end{tabularx}
}
\begin{lstlisting}
{
    "mutable": Boolean,
    "required": Boolean,
    "type": String,
    "pattern": Object,
    "default": String | Number | Boolean | Array | Object,
    "description": String
}
\end{lstlisting}

For examples of rendered Attribute Descriptions please refer, e.g., to
the Kind rendering example in Appendix~\ref{kind_instance_rendering_example}.

\section{Security Considerations}
OCCI does not require that an authentication mechanism be used nor
does it require that client to service communications are secured. It
does RECOMMEND that an authentication mechanism be used and that where
appropriate, communications are encrypted using HTTP over TLS. The
authentication mechanisms that MAY be used with OCCI are those that
can be used with HTTP and TLS. For further discussion see the
appropiate section in \cite{occi:http_protocol}.

\section{Glossary}
\label{sec:glossary}
\input{include/glossary}

\section{Contributors}
\input{include/contributors}

\section{Intellectual Property Statement}
\input{include/ip}

\section{Disclaimer}
\input{include/disclaimer}

\section{Full Copyright Notice}
\input{include/copyright}

\bibliographystyle{IEEEtran}
\bibliography{references}

\clearpage
\begin{appendices}

\lstset{
   language=json,
   basicstyle=\footnotesize,
   xleftmargin=0.0cm
}

\section{JSON Rendering Examples}
\subsection{Resouce Instance Example}
\label{resouce_instance_rendering_example}

The following is an example of a rendered \texttt{compute} resource instance as specified in section~\ref{sec:format_resource}.

\begin{lstlisting}
{
    "kind": "http://schemas.ogf.org/occi/infrastructure#compute",
    "mixins": [
        "http://example.com/occi/infrastructure/os_tpl#debian9",
        "http://example.com/occi/infrastructure/resource_tpl#medium"
    ],
    "attributes": {
        "occi.compute.speed": 2,
        "occi.compute.memory": 4,
        "occi.compute.cores": 2,
        "com.example.occi.templates.myosmixin": {
            "mykey": "myvalue"
        }
    },
    "actions": [
        "http://schemas.ogf.org/occi/infrastructure/compute/action#start"
    ],
    "id": "urn:uuid:996ad860-2a9a-504f-8861-aeafd0b2ae29",
    "links": [
        {
            "kind":"http://schemas.ogf.org/occi/infrastructure#networkinterface",
            "mixins":[
                "http://schemas.ogf.org/occi/infrastructure/networkinterface#ipnetworkinterface"
            ],
            "attributes":{
                "occi.infrastructure.networkinterface.interface":"eth0",
                "occi.infrastructure.networkinterface.mac":"00:80:41:ae:fd:7e",
                "occi.infrastructure.networkinterface.address":"192.168.0.100",
                "occi.infrastructure.networkinterface.gateway":"192.168.0.1",
                "occi.infrastructure.networkinterface.allocation":"dynamic"
            },
            "actions":[
                "http://schemas.ogf.org/occi/infrastructure/networkinterface/action#up",
                "http://schemas.ogf.org/occi/infrastructure/networkinterface/action#down"
            ],
            "id":"urn:uuid:22fe83ae-a20f-54fc-b436-cec85c94c5e8",
            "target": {
                "location:"/network/b7d55bf4-7057-5113-85c8-141871bf7635",
                "kind":"http://schemas.ogf.org/occi/infrastructure#network"
            },
            "source": {
                "location":"/compute/996ad860-2a9a-504f-8861-aeafd0b2ae29",
                "kind":"http://schemas.ogf.org/occi/infrastructure#compute"
            }
        }
    ]
}
\end{lstlisting}

\subsection{Action Invocation Example}
\label{action_invocation_rendering_example}

The following is an example of a rendered \texttt{stop} action invocation as specified in section~\ref{sec:format_action_invocation}.

\begin{lstlisting}
{
    "action":"http://schemas.ogf.org/occi/infrastructure/compute/action#stop",
    "attributes":{
        "method":"graceful"
    }
}
\end{lstlisting}

\subsection{Link Instance Example}
\label{link_instance_rendering_example}

The following is an example of a rendered \texttt{networkinterface} link as specified in section~\ref{sec:format_link}.


\begin{lstlisting}
{
    "kind":"http://schemas.ogf.org/occi/infrastructure#networkinterface",
    "mixins":[
        "http://schemas.ogf.org/occi/infrastructure/networkinterface#ipnetworkinterface"
    ],
    "attributes":{
        "occi.infrastructure.networkinterface.interface":"eth0",
        "occi.infrastructure.networkinterface.mac":"00:80:41:ae:fd:7e",
        "occi.infrastructure.networkinterface.address":"192.168.0.100",
        "occi.infrastructure.networkinterface.gateway":"192.168.0.1",
        "occi.infrastructure.networkinterface.allocation":"dynamic"
    },
    "actions":[
        "http://schemas.ogf.org/occi/infrastructure/networkinterface/action#up",
        "http://schemas.ogf.org/occi/infrastructure/networkinterface/action#down"
    ],
    "id":"urn:uuid:22fe83ae-a20f-54fc-b436-cec85c94c5e8",
    "target": {
        "location:"/network/b7d55bf4-7057-5113-85c8-141871bf7635",
        "kind":"http://schemas.ogf.org/occi/infrastructure#network"
    },
    "source": {
        "location":"/compute/996ad860-2a9a-504f-8861-aeafd0b2ae29",
        "kind":"http://schemas.ogf.org/occi/infrastructure#compute"
    }
}
\end{lstlisting}

\subsection{Kind Instance Example}
\label{kind_instance_rendering_example}

The following is an example of a rendered Kind instance as specified in section~\ref{sec:format_kind}.

\begin{lstlisting}
{
    "term":"compute",
    "scheme":"http://schemas.ogf.org/occi/infrastructure#",
    "title":"ComputeResource",
    "parent":"http://schemas.ogf.org/occi/core#resource",
    "attributes":{
        "occi.compute.hostname":{
            "mutable":true,
            "required":false,
            "type":"string",
            "description":"Hostname of the compute resource"
            "pattern": {
                "$schema": "http://json-schema.org/draft-04/schema#",
                "type": "string",
                "pattern": "\S+"
            }
        },
        "occi.compute.state":{
            "mutable":false,
            "required":false,
            "type":"string",
            "default":"inactive",
            "description":"State the compute resource is in"
        }
    },
    "actions":[
        "http://schemas.ogf.org/occi/infrastructure/compute/action#start",
        "http://schemas.ogf.org/occi/infrastructure/compute/action#stop",
        "http://schemas.ogf.org/occi/infrastructure/compute/action#restart",
        "http://schemas.ogf.org/occi/infrastructure/compute/action#suspend"
    ],
    "location":"/compute/"
}
\end{lstlisting}

\subsection{Mixin Instance Example}
\label{mixin_instance_rendering_example}

The following is an example of a rendered \texttt{medium} Resource Template Mixin as specified in section~\ref{sec:format_mixin}.

\begin{lstlisting}
{
    "term":"medium",
    "scheme":"http://example.com/template/resource#",
    "depends":[
        "http://schemas.ogf.org/occi/infrastructure#resourcetpl"
    ],
    "applies":[
        "http://schemas.ogf.org/occi/infrastructure#compute"
    ],
    "attributes":{
        "occi.compute.speed":{
            "type":"number",
            "default":2.8
        }
    },
    "title":"MediumVM",
    "location":"/template/resource/medium/"
}
\end{lstlisting}


\subsection{Action Instance Example}
\label{action_instance_rendering_example}

The following is an example of a rendered \texttt{stop} Action instance as specified in section~\ref{sec:format_action}.

\begin{lstlisting}
{
    "term":"stop",
    "scheme":"http://schemas.ogf.org/occi/infrastructure/compute/action#",
    "title":"StopComputeinstance",
    "attributes":{
        "method":{
            "mutable":true,
            "required":false,
            "type":"string",
            "default":"poweroff"
        }
    }
}
\end{lstlisting}


\section{OCCI JSON Schema}

The JSON schema provided below validates any valid OCCI message courtesy of the \texttt{anyOf} construct below. Sub-schemas or fragments need to be used to validate specific OCCI classes.

\lstset{
   language=json,
   basicstyle=\tiny,
   xleftmargin=0.0cm
}
\begin{lstlisting}
{
	"id": "http://schemas.ogf.org/occi/OCCI-schema.json",
	"$schema": "http://json-schema.org/draft-04/schema#",
	"title": "OCCI v. 1.2 JSON Rendering Schema",
	"definitions": {

		"array_of_strings": {
			"type": "array",
			"items": { "type": "string" }
		},

		"kinded_uri": {
			"id": "#kinded_uri",
			"type": "object",
			"required": ["location"],
			"additionalProperties": false,
			"properties": {
				"location": { "type": "string" },
				"kind": { "type": "string" }
			}
		},

		"resource": {
			"id": "#resource",
			"type": "object",
			"required": ["kind", "id"],
			"additionalProperties": false,
			"properties": {
				"kind": { "type": "string" },
				"mixins": { "$ref": "#/definitions/array_of_strings" },
				"attributes": { "$ref": "#/definitions/attributes" },
				"actions": { "$ref": "#/definitions/array_of_strings" },
				"id": { "type": "string" },
				"links": {
					"type": "array",
					"items": {
						"$ref": "#definitions/link"
					}
				},
				"summary": { "type": "string" },
				"title": { "type": "string" }
			}
		},

		"action_invocation": {
			"id": "#action_invocation",
			"type": "object",
			"required": ["action"],
			"additionalProperties": false,
			"properties": {
				"action": { "type": "string" },
				"attributes": { "$ref": "#/definitions/attributes" }
			}
		},

		"link": {
			"id": "#link",
			"type": "object",
			"required": ["kind", "id", "target", "source"],
			"additionalProperties": false,
			"properties": {
				"kind": { "type": "string" },
				"mixins": { "$ref": "#/definitions/array_of_strings" },
				"attributes": { "$ref": "#/definitions/attributes" },
				"actions": { "$ref": "#/definitions/array_of_strings" },
				"id": { "type": "string" },
				"source": { "$ref": "#/definitions/kinded_uri" },
				"target": { "$ref": "#/definitions/kinded_uri" },
				"rel": { "type": "string" },
				"title": { "type": "string" }
			}
		},

		"kind": {
			"id": "#kind",
			"type": "object",
			"required": ["term", "scheme"],
			"additionalProperties": false,
			"properties": {
				"term": { "type": "string" },
				"scheme": { "type": "string" },
				"title": { "type": "string" },
				"attributes": { "$ref": "#/definitions/attribute_description" },
				"actions": { "$ref": "#/definitions/array_of_strings" },
				"parent": { "type": "string" },
				"location": { "type": "string" }
			}
		},

		"mixin": {
			"id": "#mixin",
			"type": "object",
			"required": ["term", "scheme", "location"],
			"additionalProperties": false,
			"properties": {
				"term": { "type": "string" },
				"scheme": { "type": "string" },
				"title": { "type": "string" },
				"attributes": { "$ref": "#/definitions/attribute_description" },
				"actions": { "$ref": "#/definitions/array_of_strings" },
				"depends": { "$ref": "#/definitions/array_of_strings" },
				"applies": { "$ref": "#/definitions/array_of_strings" },
				"location": { "type": "string" }
			}
		},

		"action": {
			"id": "#action",
			"type": "object",
			"required": ["term", "scheme"],
			"additionalProperties": false,
			"properties": {
				"term": { "type": "string" },
				"scheme": { "type": "string" },
				"title": { "type": "string" },
				"attributes": { "$ref": "#/definitions/attribute_description" }
			}
		},

		"resource_collection": {
			"id": "#resource_collection",
			"type": "object",
			"required": ["resources"],
			"additionalProperties": false,
			"properties": {
				"resources": {
					"type": "array",
					"items": {
						"$ref": "#definitions/resource"
					}
				}
			}
		},

		"link_collection": {
			"id": "#link_collection",
			"type": "object",
			"required": ["links"],
			"additionalProperties": false,
			"properties": {
				"links": {
					"type": "array",
					"items": {
						"$ref": "#definitions/link"
					}
				}
			}
		},

		"kind_collection": {
			"id": "#kind_collection",
			"type": "object",
			"required": ["kinds"],
			"additionalProperties": false,
			"properties": {
				"kinds": {
					"type": "array",
					"items": {
						"$ref": "#definitions/kind"
					}
				}
			}
		},

		"mixin_collection": {
			"id": "#mixin_collection",
			"type": "object",
			"required": ["mixins"],
			"additionalProperties": false,
			"properties": {
				"mixins": {
					"type": "array",
					"items": {
						"$ref": "#definitions/mixin"
					}
				}
			}
		},

		"action_collection": {
			"id": "#action_collection",
			"type": "object",
			"required": ["actions"],
			"additionalProperties": false,
			"properties": {
				"actions": {
					"type": "array",
					"items": {
						"$ref": "#definitions/action"
					}
				}
			}
		},

		"attributes": {
			"id": "#attributes",
			"type": "object",
			"additionalProperties": {
				"oneOf": [
					{ "type": "number" },
					{ "type": "boolean" },
					{ "type": "string" },
					{ "type": "object" },
					{ "type": "array" }
				]
			}
		},

		"attribute_description": {
			"id": "#attribute_description",
			"oneOf": [
				{
					"type": "object",
					"additionalProperties": false,
					"minProperties": 1,
					"properties": {
						"mutable": { "type": "boolean" },
						"required": { "type": "boolean" },
						"type": { "type": "string" },
						"default": {
							"oneOf": [
								{ "type": "number" },
								{ "type": "string" },
								{ "type": "boolean" }
							]
						},
						"description": { "type": "string" },
						"pattern": { "type": "object" }
					}
				},
				{
					"type": "object",
					"additionalProperties": false,
					"patternProperties": {
						".+": {
							"$ref": "#/definitions/attribute_description"
						}
					}
				}
			]
		},

		"model": {
			"id": "#model",
			"type": "object",
			"additionalProperties": false,
			"properties": {
				"resources": {
					"type": "array",
					"items": {
						"$ref": "#definitions/resource"
					}
				},
				"links": {
					"type": "array",
					"items": {
						"$ref": "#definitions/link"
					}
				},
				"mixins": {
					"type": "array",
					"items": {
						"$ref": "#definitions/mixin"
					}
				},
				"kinds": {
					"type": "array",
					"items": {
						"$ref": "#definitions/kind"
					}
				},
				"actions": {
					"type": "array",
					"items": {
						"$ref": "#definitions/action"
					}
				}
			}
		}
	},

	"anyOf": [
		{ "$ref": "#/definitions/kind" },
		{ "$ref": "#/definitions/mixin" },
		{ "$ref": "#/definitions/action" },
		{ "$ref": "#/definitions/attributes" },
		{ "$ref": "#/definitions/attribute_description" },
		{ "$ref": "#/definitions/kind_collection" },
		{ "$ref": "#/definitions/mixin_collection" },
		{ "$ref": "#/definitions/action_collection" },
		{ "$ref": "#/definitions/resource_collection" },
		{ "$ref": "#/definitions/link_collection" },
		{ "$ref": "#/definitions/model" }
	]
}
\end{lstlisting}


\end{appendices}

\end{document}
