\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{comment}
\usepackage{occi}
%\usepackage{lineno}   % adds line numbers, may be removed for non draft versions
%\linenumbers          % adds line numbers, may be removed for non draft versions
\usepackage{verbatim} % adds verbatim options
\usepackage{tabularx} % adds extended tabular formatting options
\usepackage{listings}
\usepackage{color}
\usepackage{placeins}
\definecolor{lightgray}{rgb}{.9,.9,.9}
\definecolor{darkgray}{rgb}{.4,.4,.4}
\definecolor{purple}{rgb}{0.65, 0.12, 0.82}

\lstdefinelanguage{json}{
  ndkeywords={String, Number, Boolean, Null, Object, Array},
  ndkeywordstyle=\itshape
}
\lstset{
   language=json,
   basicstyle=\footnotesize,
}

\setlength{\headheight}{13pt}
\pagestyle{fancy}

% default sans-serif
\renewcommand{\familydefault}{\sfdefault}

% no lines for headers and footers
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% header
\fancyhf{}
\lhead{GWD-R}
\rhead{\today}

% footer
\lfoot{occi-wg@ogf.org}
\rfoot{\thepage}

% paragraphs need some space...
\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

% some space between header and text...
\headsep 13pt

\setcounter{secnumdepth}{4}

\begin{document}

% header on first page is different
\thispagestyle{empty}

GWD-R \hfill Jean Parpaillon\\
OCCI-WG \\
\rightline {September 26, 2013}\\
\rightline {Updated: \today}

\vspace*{0.5in}

\begin{Large}
\textbf{Open Cloud Computing Interface - XML Rendering}
\end{Large}

\vspace*{0.5in}

\underline{Status of this Document}

This document provides information to the community regarding the
specification of the Open Cloud Computing Interface. Distribution is
unlimited.

\underline{Copyright Notice}

Copyright \copyright Open Grid Forum (2013). All Rights Reserved.

\underline{Trademarks}

OCCI is a trademark of the Open Grid Forum.

\underline{Abstract}

This document, part of a document series, produced by the OCCI working
group within the Open Grid Forum (OGF), provides a high-level
definition of a Protocol and API. The document is based upon
previously gathered requirements and focuses on the scope of important
capabilities required to support modern service offerings.

\underline{Comments}
\newcommand{\ralf}[1]{\textcolor{red}{RN: #1}}
\newcommand{\andy}[1]{\textcolor{green}{AE: #1}}
\newcommand{\florian}[1]{\textcolor{blue}{FF: #1}}
\newcommand{\jean}[1]{\textcolor{purple}{JP: #1}}

\newpage
\tableofcontents
\newpage

\section{Introduction}
\input{include/introduction}

\section{Notational Conventions}
\input{include/notational}

\section{OCCI XML Rendering}

The OCCI XML Rendering specifies a rendering of OCCI instance types in
the eXtensible Markup Language (XML), as defined in \cite{w3c:xml11}.

The Rendering can be used to render OCCI instances independently of the
transport mechanism being used. Thus messages can be delivered by e.g. the HTTP
protocol as specified in \cite{occi:http_rendering} or by using text files with
the .xml file extension as defined in \cite{w3c:xml11}.

The following media-type MUST be used for the OCCI XML Rendering:

{\tt application/occi+xml}

\section{XML Format}
\label{sec:xml_format}

The OCCI XML Rendering consists of a sequence of XML element containing
information on the OCCI Core instances OCCI Kind, OCCI Mixin, OCCI
Action, OCCI Link and OCCI Resource. The rendering also include an XML
element to invoke the operation identified by OCCI Actions.
The rendering of each OCCI Core instance will be described in the
following sections.

\jean{Planned, not present in XML schema}

OCCI XML rendering include XML element to represent the
CRUD\footnote{Create, Retrieve, Update, Delete} actions on OCCI types.

\subsection{Category Format}
\label{sec:format_category}
The OCCI Category Format consists of a XML object.
Table~\ref{tbl:format_category} defines the object members.

\begin{lstlisting}
<xs:complexType name="categoryType">	
	<xs:attributeGroup ref="occi:categoryGroup"></xs:attributeGroup>
</xs:complexType>

<xs:attributeGroup name="categoryIdGroup" >
	<xs:attribute name="scheme" type="xs:anyURI"></xs:attribute>
	<xs:attribute name="term" type="xs:string" use="required"></xs:attribute>
</xs:attributeGroup>

<xs:attributeGroup name="categoryGroup">	
	<xs:attributeGroup ref="occi:categoryIdGroup"></xs:attributeGroup>
	<xs:attribute name="title" type="xs:string"></xs:attribute>
	<xs:attribute name="use">
		<xs:simpleType>
			<xs:restriction base="xs:string">
				<xs:enumeration value="required"></xs:enumeration>
				<xs:enumeration value="optional"></xs:enumeration>
			</xs:restriction>
		</xs:simpleType>
	</xs:attribute>
</xs:attributeGroup>
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_category}
	Attribute defined for the Categorie type}
	{
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Attribute & XML type & Description & Required \\
    \colrule
    term & xs:string & Unique identifier of the Category instance within the categorisation scheme & yes \\
    scheme & xs:anyURI & The categorisation scheme & yes \\
    title & xs:string & The display name of an instance & no \\
    use & required & Whether the OCCI Attribute must be supplied by the client at instance creation time & no \\
    \botrule
    \end{tabularx}
}
\FloatBarrier

\subsection{Attribute Description Format}
\label{sec:format_attribute_description}

OCCI Attribute Descriptions are rendered as XML objects as defined in table~\ref{tbl:format_attribute_description} and table~\ref{tbl:format_attribute_description_Att}.

\begin{lstlisting}
<xs:complexType name="attributeSpecType">
	<xs:complexContent>
		<xs:extension base="xs:attribute">
			<xs:attribute name="immutable" type="xs:boolean" default="false">
				<xs:annotation>
					<xs:documentation>
						If 'immutable' is true, the attribute can
						not be
						modified by user.
					</xs:documentation>
				</xs:annotation>
			</xs:attribute>
		</xs:extension>
	</xs:complexContent>
</xs:complexType>

<xs:complexType name="attributeType">
	<xs:sequence>
		<xs:element name="value" type="xs:anyURI" minOccurs="0" maxOccurs="unbounded" />
	</xs:sequence>
	<xs:attribute name="name" type="xs:string"></xs:attribute>
	<xs:attribute name="value" type="xs:string"></xs:attribute>
</xs:complexType>
\end{lstlisting}


\mytablefloat{
    \label{tbl:format_attribute_description}
    Element defined for the Attribute type
    } {
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Element & XML type & Description & Mutability & Multiplicity \\
    \colrule
    value & xs:anyURI & Notion of array & Immutable & 0..* \\
    \botrule
    \end{tabularx}
}
\mytablefloat{	   
    \label{tbl:format_attribute_description_Att}
     Attribute defined for the Attribute type
     }{
     \begin{tabularx}{\textwidth}{llXll}
    \toprule
     Attribute & XML type & Description & Required \\
    \colrule
    name & xs:string & OCCI Attribute name & yes\\
    value & xs:string & Description & no\\
    type & XML type & OCCI Attribute type & no\\
    title & String & A descriptin of the OCCI Attribute & no \\
    use & required & Whether the OCCI Attribute must be supplied by the client at instance creation time & no\\
    default & XML type & OCCI Attribute default value & no\\
    immutable & xs:boolean & OCCI Attribute mutability & no \\
    \botrule
    \end{tabularx}
}
\FloatBarrier

\begin{lstlisting}
Example:
<occi:attribute name="occi.compute.state" use="required"
		default="inactive" immutable="true"
		title="System state" type="xs:string" />
\end{lstlisting}

\subsection{Entity Format}
\label{sec:format_entity}

The OCCI Entity Format consists of a XML object.
Table~\ref{tbl:format_entity} and table~\ref{tbl:format_entity_Att} define the object members.

\begin{lstlisting}
<xs:complexType name="entityType">
	<xs:sequence>
		<xs:element name="kind" type="occi:categoryType" minOccurs="1" maxOccurs="1" />
		<xs:element name="mixin" type="occi:categoryType" minOccurs="0" maxOccurs="unbounded" />
		<xs:element name="attribute" type="occi:attributeType" minOccurs="0" maxOccurs="unbounded" />
		<xs:element name="action" type="occi:categoryType" minOccurs="0" maxOccurs="unbounded" />
	</xs:sequence>
	<xs:attribute name="id" type="xs:anyURI" ></xs:attribute>
	<xs:attribute name="title" type="xs:string" ></xs:attribute>
	<xs:attribute name="href" type="xl:href" />
</xs:complexType>
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_entity}
    Element defined for the Entity type
    } {
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Element & XML type & Description & Mutability & Multiplicity \\
    \colrule
    kind & occi:categoryType & Type identifier & immutable & 1 \\

    mixins & occi:categoryType & List of type identifiers of associated OCCI Mixins & mutable & 0..* \\

    attributes & occi:attributeType & Instance Attributes (see
\ref{sec:format_attribute_description}) & mutable & 0..* \\
    
    actions & occi:category & List of type identifiers of OCCI
Actions applicable to the OCCI Resource instance & mutable & 0..* \\
    \botrule
    \end{tabularx}
}
\mytablefloat{	   
    \label{tbl:format_entity_Att}
     Attribute defined for the Entity type
     }{    
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Attribute & XML type & Description & Required \\
    \colrule
    id & xs:anyURI & ID of the OCCI Resource & yes\\

    title & xs:string & A Description of the OCCI Entity  & yes\\
            
    href & xl:href & Path of the OCCI Entity & yes\\
    \botrule
    \end{tabularx}
}
\FloatBarrier

\subsection{Resource Instance Format}
\label{sec:format_resource}

The OCCI Resource Instance Format consists of a XML object as shown in the
following example. Section \ref{sec:example_resource} contains a detailed
example.
Table~\ref{tbl:format_resource} and table~\ref{tbl:format_resource_Att} define the object members.

\begin{lstlisting}
<xs:complexType name="resourceType">
	<xs:complexContent>
		<xs:extension base="occi:entityType">
			<xs:sequence>
				<xs:element name="summary" type="xs:string" minOccurs="0" maxOccurs="1" />
				<xs:element name="link" type="occi:resourceLinkType" minOccurs="0" maxOccurs="unbounded" />
			</xs:sequence>
		</xs:extension>
	</xs:complexContent>
</xs:complexType>
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_resource}
    OCCI Resource instances are rendered inside the top-level XML object with
name
{\em resources}
    as a ressourceType of XML objects with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Element member & XML type & Description & Mutability & Multiplicity \\
    \colrule
    entity & occi:entityType & Set of entity instances & Mutable & 0..* \\
    \botrule
    \end{tabularx}
}
\mytablefloat{	   
    \label{tbl:format_resource_Att}
     Attribute defined for the Entity type
     }{      
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Attribute & XML type & Description & Required \\
    \colrule
    summary & xs:string & A summarising description of the Ressource instance & no\\
            
    links & occi:resourceLinkType & List of URIs of OCCI Links & no\\
    \botrule
    \end{tabularx}
}
\FloatBarrier

\begin{lstlisting}
Example:
<resources id="xs:anyURI" title="xs:string" href="xl:href">
	<kind type="occi:categoryType"/>
	<mixins type="occi:categoryType"/>
        <attributes type="occi:attributeType"/>
        <actions: type="occi:attributeType"/>
        <links type="occi:resourceLinkType"> 
</resources>
\end{lstlisting}

\subsection{Link Instance Format}
\label{sec:format_link}

The OCCI Link Instance Format consists of a XML object as shown in the
following example. Section \ref{sec:example_link} contains a detailed example.
Table~\ref{tbl:format_link} defines the object members.

\begin{lstlisting}
<xs:complexType name="linkType">
	<xs:complexContent>
		<xs:extension base="occi:entityType" />
	</xs:complexContent>
</xs:complexType>
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_link}
    OCCI Resource instances are rendered inside the top-level XML object with
name
{\em resources}
    as a linkType of XML objects with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Element & XML type & Description & Mutability & Multiplicity \\
    \colrule
    entity & occi:entityType & Set of entity instances & Mutable & 0..* \\
    \botrule
    \end{tabularx}
}
\FloatBarrier

\begin{lstlisting}
Example:
<link id="xs:anyURI" title="xs:string" href="xl:href">
	<kind type="occi:categoryType"/>
	<mixins type="occi:categoryType"/>
        <attributes type="occi:attributeType"/>
        <actions: type="occi:attributeType"/>
        <links type="occi:resourceLinkType">
<link>
\end{lstlisting}

\subsection{Kind Instance Format}
\label{sec:format_kind}

The OCCI Kind Instance Format consists of a XML object as shown in the
following example. Section \ref{sec:example_kind} contains a detailed example.
Table~\ref{tbl:format_kind} and table~\ref{tbl:format_kind_Att} define the top-level object members.

\begin{lstlisting}
<xs:complexType name="kindType">
	<xs:sequence>
		<xs:element name="parent" type="occi:categoryType" minOccurs="1" maxOccurs="1" />
		<xs:element name="attribute" type="occi:attributeSpecType" minOccurs="0" maxOccurs="unbounded" />
		<xs:element name="action" type="occi:actionSpecType" minOccurs="0" maxOccurs="unbounded" />
	</xs:sequence>
	<xs:attributeGroup ref="occi:categoryGroup"></xs:attributeGroup>
	<xs:attribute name="location" type="xs:anyURI"></xs:attribute>
</xs:complexType>
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_kind}
    OCCI Kind instances are rendered inside the top-level XML object with name
{\em kinds} as a kindType of XML objects with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Element & XML type & Description & Mutability & Multiplicity \\
    \colrule
    parent & occi:categoryType & OCCI Kind type identifier of the
related ``parent'' \hl{Kind} instance & immutable & 1 \\

    attributes & occi:attributeSpecType & Attribute description & immutable & 0..* \\

    actions & occi:actionSpecType & List of OCCI Action type
identifiers & immutable & 0..* \\
    \botrule
    \end{tabularx}
}
\mytablefloat{	   
    \label{tbl:format_kind_Att}
     Attribute defined for the Entity type
     }{     
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Attribute & XML type & Description & Required \\
    \colrule
    location & xs:anyURI & Transport protocol specific URI bound to the OCCI Kind
instance. MUST be supplied for the OCCI Kinds of all OCCI Entities except OCCI
Entity itself & no \\
   
    attributeGroup & occi:attributeGroup & Set of category & yes \\
    \botrule
    \end{tabularx}
}
\FloatBarrier

\begin{lstlisting}
Example:
<kinds location="xs:anyURI">
	<parent type="occi:categoryType"/>
	<attribute type="occi:attributeSpecType"/>
	<action type="occi:actionSpecType"/>
</kind>
\end{lstlisting}

\subsection{Mixin Instance Format}
\label{sec:format_mixin}

The OCCI Mixin Instance Format consists of a XML object as shown in the following example. Section \ref{sec:example_mixin} contains a detailed example.
Table~\ref{tbl:format_mixin} and table~\ref{tbl:format_mixin} define the top-level object members.
 
\begin{lstlisting}
<xs:complexType name="mixinType">
	<xs:sequence>
		<xs:element name="depends" type="occi:categoryType" minOccurs="0" maxOccurs="unbounded" />
		<xs:element name="applies" type="occi:categoryType" minOccurs="0" maxOccurs="unbounded" />
		<xs:element name="attribute" type="occi:attributeSpecType" minOccurs="0" maxOccurs="unbounded" />
		<xs:element name="action" type="occi:actionSpecType" minOccurs="0" maxOccurs="unbounded" />
	</xs:sequence>
	<xs:attributeGroup ref="occi:categoryGroup"></xs:attributeGroup>
	<xs:attribute name="location" type="xs:anyURI"></xs:attribute>
</xs:complexType>
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_mixin}
    OCCI Mixin instances are rendered inside the top-level XML object with name
{\em mixins} as a mixinType of XML objects with the following entries:
    } {
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Element & XML type & Description & Mutability & Multiplicity \\
    \colrule
    attributes & occi:attributeSpecType & Attribute description, see
~\ref{tbl:format_attribute_description} & immutable & 0..* \\

    depends & occi:categoryType & List of type identifiers of the dependent
 \hl{Mixin} instances & immutable & 0..* \\
 
    applies & occi:categoryType & List of OCCI Kind type identifiers this OCCI 
Mixin can be applied to \\

    actions & occi:actionSpecType & List of OCCI Action type identifiers
& immutable & 0..* \\
    \botrule
    \end{tabularx}
}
\mytablefloat{	   
    \label{tbl:format_mixin_Att}
     Attribute defined for the Entity type
     }{  
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Attribute & XML type & Description & Required \\
    \colrule
    location & xs:anyURI & Transport protocol specific URI bound to the OCCI Mixin
instance & no \\

    attributeGroup & occi:attributeGroup & Set of category & yes \\
    \botrule
    \end{tabularx}
}
\FloatBarrier

\begin{lstlisting}
Example:
<mixins location="xs:anyURI">
	<depends type="occi:categoryType"/>
	<applies type="occi:categoryType"/>
	<attribute type="occi:attributeSpecType"/>
	<action type="occi:actionSpecType"/>
</mixins>
\end{lstlisting}

\subsection{Action Instance Format}
\label{sec:format_action}

The OCCI Action Instance Format consists of a XML object as shown in the
following example.
Table~\ref{tbl:format_action} and table~\ref{tbl:format_action_Att} define the top-level object members.

\begin{lstlisting}
<xs:complexType name="actionSpecType">
	<xs:sequence>
		<xs:element name="attribute" type="occi:attributeSpecType" minOccurs="0" maxOccurs="unbounded"/>
	</xs:sequence>
	<xs:attributeGroup ref="occi:categoryGroup"></xs:attributeGroup>
</xs:complexType>
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_action}
    An OCCI Action invocation is rendered as top-level XML object with
 the following entries:
    } {
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Element & XML type & Description & Mutability & Multiplicity \\
    \colrule
    
    attributes & occi:attributeSpecType & Instance attributes & mutable & 0..* \\ 
    \botrule
    \end{tabularx}
}
\mytablefloat{	   
    \label{tbl:format_action_Att}
     Attribute defined for the Entity type
     }{      
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Attribute & XML type & Description & Required \\
    \colrule
    attributeGroup & occi:attributeGroup & Set of category & yes\\
    \botrule
    \end{tabularx}
}
\FloatBarrier

\begin{lstlisting}
Example:
<occi:action term="suspend"
scheme="http://schemas.ogf.org/occi/infrastructure/compute/action#"
title="Suspend the system (hibernate or in RAM)" >
	<occi:attribute name="method" type="xs:string" />
</occi:action>
\end{lstlisting}

\subsection{Action Invocation Format}
\label{sec:format_action_invocation}

The OCCI Action Invocation  Format identifies an invocable operation on a OCCI Resource or
OCCI Link instance. To trigger such an operation the OCCI Action Invocation
Format is required.

The OCCI Action Invocation Format consists of a top-level XML object as shown in the
following example. Section \ref{sec:example_action_invocation} contains a detailed example.
Table~\ref{tbl:format_action_invocation} and table~\ref{tbl:format_action_invocation_Att} define the object members.

\begin{lstlisting}
<xs:complexType name="actionType">
	<xs:sequence>
		<xs:element name="attribute" type="occi:attributeType"></xs:element>
	</xs:sequence>
	<xs:attributeGroup ref="occi:categoryGroup"></xs:attributeGroup>
</xs:complexType>
\end{lstlisting}

\mytablefloat{
    \label{tbl:format_action_invocation}
    An OCCI Action invocation is rendered as top-level XML object with
 the following entries:
    } {
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Element & XML type & Description & Mutability & Multiplicity \\
    \colrule
    
    attribute & occi:attributeType & Instance attributes & mutable & 0..* \\
    \botrule
    \end{tabularx}
}
\mytablefloat{	   
    \label{tbl:format_action_invocation_Att}
     Attribute defined for the Entity type
     }{      
    \begin{tabularx}{\textwidth}{llXll}
    \toprule
    Attribute & XML type & Description & Required \\
    \colrule
    attributeGroup & occi:categoryGroup & Set of category & yes \\
    
    \botrule
    \end{tabularx}
}
\FloatBarrier

\begin{lstlisting}
Example:
<action xmlns="http://schemas.ogf.org/occi"
term="stop" scheme="http://schemas.ogf.org/occi/infrastructure/compute/action#" >
	<attribute name="method" value="graceful" />
</action>
\end{lstlisting}

\section{Appendix: Full XML schema}
\label{sec:full_xml_schema}

\subsection{OCCI XML schema}
\label{sec:occi_xml_schema}

\lstinputlisting[language=XML]{schemas/occi.xsd}

\section{Detailed Examples}
\label{sec:examples}

%FIXME: Check that examples are correct and cover most aspects

\subsection{Resource Instance Format Example}
\label{sec:example_resource}

\begin{lstlisting}
<resource xmlns='http://schemas.ogf.org/occi' xmlns:xl='http://www.w3.org/2008/06/xlink' title='Machine a toto'>
	<kind scheme='http://schemas.ogf.org/occi/infrastructure#'term='compute'/>
	<attribute name='occi.compute.speed' value='4.0e3'/>
	<attribute name='occi.compute.memory' value='5.0'/>
	<attribute name='occi.compute.hostname' value='pc_toto'/>
	<attribute name='occi.compute.cores' value='2'/>
	<attribute name='occi.compute.architecture' value='x86'/>
	<link id='xmpp+occi:user-1@localhost/store/mylinks/json/networkinterface/0008413e-4091-42a8-9e01-f93c8c40ec21'>
	<kind scheme='http://schemas.ogf.org/occi/infrastructure#'term='networkinterface'/>
		<attribute name='occi.core.target' xl:href='xmpp+occi:user-1@localhost/store/myresources/xml/network/6a7d9971-214d-44c1-9da2-af1229584844'/>
		<attribute name='occi.networkinterface.address' value='192.168.3.4'/>
		<attribute name='occi.networkinterface.allocation' value='dynamic'/>
		<attribute name='occi.networkinterface.gateway' value='192.168.3.0'/>
		<attribute name='occi.networkinterface.interface' value='eth0'/>
		<attribute name='occi.networkinterface.mac' value='00:80:41:ae:fd:32'/>
		<mixin scheme='http://schemas.ogf.org/occi/infrastructure/networkinterface#' term='ipnetworkinterface'/>
	</link>
</resource>
\end{lstlisting}

\subsection{Link Instance Format Example}
\label{sec:example_link}

\begin{lstlisting}
<link xmlns='http://schemas.ogf.org/occi' xmlns:xl='http://www.w3.org/2008/06/xlink' id='xmpp+occi:user-1@localhost/store/mylinks/xml/networkinterfaces/id9e295664-afff-419b-8a1d-e092bdf6f718'>
	<attribute name='occi.core.source' xl:href='xmpp+occi:user-1@localhost/store/myresources
	/xml/compute/f322e40a-6c9f-42b0-8a63-231f7ac7f3e4'/>
	<kind scheme='http://schemas.ogf.org/occi/infrastructure#' term='networkinterface'/>
	<attribute name='occi.core.target' xl:href='xmpp+occi:user-1@localhost/store/myresources/xml/network/6a7d9971-214d-44c1-9da2-af1229584844'/>
	<attribute name='occi.networkinterface.address' value='192.168.3.4'/>
	<attribute name='occi.networkinterface.allocation' value='dynamic'/>
	<attribute name='occi.networkinterface.gateway' value='192.168.3.0'/>
	<attribute name='occi.networkinterface.interface' value='eth0'/>
	<attribute name='occi.networkinterface.mac' value='00:80:41:ae:fd:32'/>
	<mixin scheme='http://schemas.ogf.org/occi/infrastructure/networkinterface#' term='ipnetworkinterface'/>
</link>
\end{lstlisting}

\subsection{Action Invocation Format Example}
\label{sec:example_action_invocation}

\begin{lstlisting}
<occi:action xmlns="http://schemas.ogf.org/occi"
term="stop" scheme="http://schemas.ogf.org/occi/infrastructure/compute/action#" >
	<occi:attribute name="method" value="graceful"/>
</occi:action>
\end{lstlisting}

\subsection{Kind Format Example}
\label{sec:example_kind}
\begin{lstlisting}
<occi:kind scheme="http://schemas.ogf.org/occi/infrastructure#" term="compute" title="Compute Ressource">
	<occi:attribute name="occi.compute.hostname.mutable" value="true"/>
	<occi:attribute name="occi.compute.hostname.required" value="false"/>
	<occi:attribute name="occi.compute.hostname.type" value="xs:string"/>
	<occi:attribute name="occi.compute.hostname.description" value="Hostname of the compute ressource"/>
	<occi:attribute name="occi.compute.state.mutable" value="false"/>
	<occi:attribute name="occi.compute.hostname.required" value="false"/>
	<occi:attribute name="occi.compute.hostname.type" value="xs:string"/>
	<occi:attribute name="occi.compute.hostname.default" value="inactive"/>
	<occi:attribute name="occi.compute.hostname.description" value="State the compute ressource is in"/>
	<occi:action sechme="http://schemas.ogf.org/occi/infrastructure/compute/action#start"/>
	<occi:action sechme="http://schemas.ogf.org/occi/infrastructure/compute/action#stop"/>
	<occi:action sechme="http://schemas.ogf.org/occi/infrastructure/compute/action#restart"/>
	<occi:action sechme="http://schemas.ogf.org/occi/infrastructure/compute/action#suspend"/>
</occi:kind>
\end{lstlisting}

\subsection{Mixin Format Example}
\label{sec:example_mixin}

\begin{lstlisting}
<occi:mixin scheme="http://schemas.ogf.org/occi/infrastructure#" term="ressource_tpl" title="Medium VM">
	<occi:attribute name="occi.compute.speed.type" value="number"/>
	<occi:attribute name="occi.compute.speed.default" value="2.8">		
</occi:mixin>
\end{lstlisting}

\subsection{Action Format Example}
\label{sec:example_action} 

\begin{lstlisting}
<occi:action scheme="http://schemas.ogf.org/occi/infrastructure/compute/action#" term="stop" title="Stop Compute instance">
	<occi:attribute name="method.mutable" value="true"/>
	<occi:attribute name="method.required" value="false"/>
	<occi:attribute name="method.type" value="string"/>
	<occi:attribute name="method.default" value="poweroff"/>
</occi:action>
\end{lstlisting}

\section{Glossary}
\label{sec:glossary}
\input{include/glossary}

\section{Contributors}
\include{include/contributors}
% FIXME: Insert an up-to-date table here!

\section{Intellectual Property Statement}
\input{include/ip}

\section{Disclaimer}
\input{include/disclaimer}

\section{Full Copyright Notice}
\input{include/copyright}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
