The Open Cloud Computing Interface (OCCI) is a RESTful Protocol and
API for all kinds of management tasks. OCCI was originally initiated
to create a remote management API for IaaS%
\footnote{Infrastructure as a Service}
model-based services, allowing for the development of interoperable tools for
common tasks including deployment, autonomic scaling and monitoring.
%
It has since evolved into a flexible API with a strong focus on
interoperability while still offering a high degree of extensibility. The
current release of the Open Cloud Computing Interface is suitable to serve many
other models in addition to IaaS, including PaaS and SaaS.

In order to be modular and extensible the current OCCI specification is
released as a suite of complementary documents, which together form the complete
specification.
%
The documents are divided into four categories consisting of the OCCI Core,
the OCCI Protocols, the OCCI Renderings and the OCCI Extensions.
%
\begin{itemize}
\item The OCCI Core specification consists of a single document defining the
 OCCI Core Model. The OCCI Core Model can be interacted with through {\em
 renderings} (including associated behaviors) and expanded through {\em extensions}.
\item The OCCI Protocol specifications consist of multiple documents, each
 describing how the model can be interacted with over a particular protocol (e.g. HTTP, AMQP, etc.).
 Multiple protocols can interact with the same instance of the OCCI Core Model.
\item The OCCI Rendering specifications consist of multiple documents, each
 describing a particular rendering of the OCCI Core Model. Multiple renderings can
 interact with the same instance of the OCCI Core Model and will automatically support
 any additions to the model which follow the extension rules defined in OCCI
 Core.
\item The OCCI Extension specifications consist of multiple documents,
  each describing a particular extension of the OCCI Core Model. The
  extension documents describe additions to the OCCI Core Model
  defined within the OCCI specification suite.
\end{itemize}
%

The current specification consists of seven documents. This
specification describes version 1.2 of OCCI and is backward compatible with 1.1.
Future releases of OCCI
may include additional protocol, rendering and extension specifications. The specifications to be
implemented (MUST, SHOULD, MAY) are detailed in the table below.

\mytablefloat{
	\label{tbl:occi_compliancy}%
	What OCCI specifications must be implemented for the specific version.
}
{
	\begin{tabular}{lll}
	\toprule
	Document & OCCI 1.1 & OCCI 1.2 \\
	\colrule
	Core Model & MUST & MUST \\
	Infrastructure Model  & SHOULD & SHOULD \\
	Platform Model & MAY & MAY \\
	SLA Model & MAY & MAY \\
	HTTP Protocol & MUST & MUST \\
	Text Rendering& MUST & MUST \\
	JSON Rendering& MAY & MUST \\
	\botrule
	\end{tabular}
}

% hello
